---
title: "Matt's Notes on Plotly"
date: '2022-09-15'
output:
  html_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

From <https://plotly.com/r>.  Load with `library(plotly)`.  

```{r,include=FALSE}
library(plotly)
```

## Figures
Main function is `plot_ly`, used to generate plotly figure objects

```{r}
fig <- plot_ly(midwest, x = ~percollege, color = ~state, type = "box")
fig
```

Note that `fig` is a ploty object, which is a sub-class of htmlwidget.  

```{r}
class(fig)
```

Plotly represents figures as nested lists of attributes, which enables it to be represented as JSON and passed to plotly.js library. To make this clearer, build new figure with several attributes:

```{r}
fig <- plot_ly() %>% 
  add_lines(x = c("a","b","c"), y = c(1,3,2))%>% 
  layout(title="sample figure", xaxis = list(title = 'x'), yaxis = list(title = 'y'), plot_bgcolor = "#c7daec") 
```

In this example, the layout attribute has sub-attributes of title, xaxis, etc.  And xaxis attribute has sub-attribute of title, and this often continues through more nested named lists.  

All plotly figures have three top-level attributes:

* data
* layout
* frames
$\sqrt(\pi)$

F


