---
title: "Ch28_R_Markdown_Slideshow"
author: "Matt Smith"
date: "1/5/2021"
output: ioslides_presentation

---

Two common formats are ioslides (`output: ioslides_presentation` in yaml headers), and revealjs (`output: revealjs::revealjs_presentation`).  Since the revealjs one comes from revealjs package and not the rmarkdown package, we need to specify the reavealjs package.  

Slides are specified by double hashtags (##).  

The resulting slideshow can embed interactive graphics, though on first glance, it appears quite tricky to get the content to fit as desired.  

## First Slide

A list of things to cover

- First Item

- Second Item

- Third Item



## Some R Code

The code below will generate some results and a plot.

```{r pressure, echo=FALSE}
plot(pressure)
```


## Another Slide

Some more information goes here



## Some Links

[My Website](http://www.jaredlander.com)



[R Bloggers](http://www.r-bloggers.com)

## datatables

Use the `DT` package to generate interactive displays of tabular data.

```{r DT}
#install.packages('DT')
library(DT)
data(diamonds, package='ggplot2')
#DT package load the datatable function
datatable(head(diamonds, 100))
```


